State of the Practice for Medical Imaging Software Based on Open Source Repositories

### 2024-11-15 StateOfPractice/Papers/SOP_MI_CritReviewsBioMedEng/SOP_MI_OSR.tex
Revisions to conclusions
Submitted version of paper, letter to editor and copyright transfer

We present the state of the practice for the development of Medical Imaging (MI)
software based on data available in open source repositories. We selected 29
projects from 48 candidates and assessed 9 software qualities (installability,
correctness/verifiability, reliability, robustness, usability, maintainability,
reusability, understandability, and visibility/transparency) by answering 108
questions for each. Using the Analytic Hierarchy Process (AHP) on the
quantitative data, we ranked the MI software.  The top five are \textit{3D
Slicer}, \textit{ImageJ}, \textit{Fiji}, \textit{OHIF Viewer}, and
\textit{ParaView}.  This is consistent with the community's view, with four of
these also appearing in the top five using GitHub metrics (stars-per-year).  The
quality and quantity of documentation present in a project correlates quite well
with its popularity.  Generally, MI software is in a healthy state: in the
repositories we observed 88\% of the documentation artifacts recommended by
research software development guidelines, 100\% of MI projects use version
control tools, and developers appear to use the common quasi-agile research
software development process. However, the current state of the practice
deviates from existing guidelines as some recommended artifacts are rarely
present (like a test plan, requirements specification, code of conduct, and code
style guidelines), low usage of continuous integration (17\% of the projects),
low use of unit testing (about 50\% of projects), and room for improvement with
documentation. From developer interviews, we identified 7 concerns: lack of
development time, lack of funding, technology hurdles, correctness, usability,
maintainability, and reproducibility. We recommend: increase effort on
documentation, increase testing by enriching datasets, increase continuous
integration usage, move to web applications, employ linters, use peer reviews,
and design for change.

@@

While we did find a reasonable amount of documentation, especially when consider
all MI projects, there were definitely some holes. Some important documentation
(test plans and requirements documentation) was missing, and other
(contributors' guide, code of conduct, code style guidelines, product roadmap,
design documentation, and API documentation) were rare.

>>>>

We review the state of the practice for the development of Medical Imaging (MI)
software based on data available in open-source repositories. We selected 29
projects from 48 candidates and assessed 9 software qualities by answering 108
questions for each. Using the Analytic Hierarchy Process (AHP) on the
quantitative data, we ranked the MI software.  The top five are \textit{3D
Slicer}, \textit{ImageJ}, \textit{Fiji}, \textit{OHIF Viewer}, and
\textit{ParaView}.  This is consistent with the community's view, with four of
these also appearing in the top five using GitHub metrics (stars-per-year).  The
quality and quantity of documentation present in a project correlate quite well
with its popularity.  Generally, MI software is in a healthy state: in the
repositories, we observed 88\% of the documentation artifacts recommended by
research software development guidelines and 100\% of MI projects use version
control tools. However, the current state of the practice
deviates from existing guidelines as some recommended artifacts are rarely
present (like a test plan, requirements specification, and code
style guidelines), low usage of continuous integration (17\% of the projects),
low use of unit testing (about 50\% of projects), and room for improvement with
documentation. From developer interviews, we identified 7 concerns: lack of
development time, lack of funding, technology hurdles, correctness, usability,
maintainability, and reproducibility. We recommend: increasing effort on
documentation, increasing testing by enriching datasets, increasing continuous
integration, moving to web applications, employing linters, using peer reviews,
and designing for change.

@@

While we did find a reasonable amount of documentation, especially when we
consider all MI projects, there were definitely some holes. Some important
documentation (test plans and requirements documentation) was missing, and other
(contributors' guide, code of conduct, code style guidelines, product roadmap,
design documentation, and API documentation) were rare.
